---
title: "CP1"
author: "Brian Liswell"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Repro Course Project 1

### Commit containing full submission

### 1. Code for reading in the dataset and/or processing the data

activity.csv contains 3 variables, steps, date and interval.  Each 5 second interval ranges from 0 to 2355 at the end of the day.

```{r }
data <- read.csv('activity.csv')
summary(data)
```

### 2. Histogram of the total number of steps taken each day


```{r echo = TRUE}
data_by_date <- split(data,data$date)

date_names <- names(data_by_date)

steps_per_day <-numeric(length(date_names))
for(i in 1:length(steps_per_day)){  #length(steps_per_day)
        temp<-data_by_date[[date_names[i]]]$steps
        steps_per_day[i] <- sum(temp)
        
}
hist(steps_per_day)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


### 3. Mean and median number of steps taken each day
```{r echo = TRUE}
mean(steps_per_day,na.rm = TRUE)
```
```{r echo = TRUE}
median(steps_per_day,na.rm = TRUE)
```

### 4. Time series plot of the average number of steps taken.
I think this means the total number of steps taken per day.  Otherwise, over what time period is the "average"

```{r echo = TRUE}
library(ggplot2)

data_total <-data.frame(day = as.Date(date_names),value = steps_per_day)

p2 <- ggplot(data_total, aes(x = day, y = steps_per_day)) + 
        geom_line() +
        xlab("")
p2
```

### OR
4. means to average accross all the 5 minute intervals from each day, which could be how to interpret number 5.
```{r echo = TRUE}
steps_per_5min <-numeric(length(data_by_date[[date_names[1]]]$steps))
interval <- data_by_date[[date_names[1]]]$interval
cnt <-0
for(i in 1:length(date_names)){  #length(steps_per_day)
        temp<-data_by_date[[date_names[i]]]$steps
        temp[is.na(temp)] <-0
        steps_per_5min <- steps_per_5min + temp
        cnt <- cnt+1
}
avg_per_5min <- steps_per_5min/cnt
interval_avg <-data.frame(interval = interval,value = avg_per_5min)

p3 <- ggplot(interval_avg, aes(x = interval, y = value)) + 
        geom_line() +
        xlab("interval value, seconds")+
        ylab("Average number of steps accross all days")
p3
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r echo = TRUE}
max_ind <- which.max(steps_per_5min)
print(paste0("interval number ",max_ind))

print(paste0("At this many seconds into the day ",interval[max_ind]))
```

### 6. Code to describe and show a strategy for imputing missing data

I did this above in step 4, with temp[is.na(temp)] <-0, which replaces the missing data with 0.  I did not do this for the code in step 2.  See the effects in step 7.

### 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r echo = TRUE}
data_by_date <- split(data,data$date)

date_names <- names(data_by_date)

steps_per_day <-numeric(length(date_names))
for(i in 1:length(steps_per_day)){  #length(steps_per_day)
        temp<-data_by_date[[date_names[i]]]$steps
        temp[is.na(temp)] <-0
        steps_per_day[i] <- sum(temp)
        
}
hist(steps_per_day)
```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

"What is a Week 'end'?"
```{r echo = TRUE}
date_names_Date <- as.Date(date_names)
library(timeDate)



wd_steps_per_5min <-numeric(length(data_by_date[[date_names[1]]]$steps))
we_steps_per_5min <-numeric(length(data_by_date[[date_names[1]]]$steps))
interval <- data_by_date[[date_names[1]]]$interval
wd_cnt <-0
we_cnt <-0

for(i in 1:length(date_names)){  #length(steps_per_day)
        temp<-data_by_date[[date_names[i]]]$steps
        temp[is.na(temp)] <-0
        
        if(isWeekday(date_names_Date[i], wday=1:5)){
                wd_steps_per_5min <- wd_steps_per_5min + temp
                wd_cnt <- wd_cnt+1
                
        }
        else{
                we_steps_per_5min <- we_steps_per_5min + temp
                we_cnt <- wd_cnt+1
        }
}
wd_avg_per_5min <- wd_steps_per_5min/wd_cnt
we_avg_per_5min <- we_steps_per_5min/we_cnt
compare_avg <-data.frame(interval = interval,wd = wd_avg_per_5min,we = we_avg_per_5min)
we_avg <-data.frame(interval = interval,value = we_avg_per_5min)
wd_avg <-data.frame(interval = interval,value = wd_avg_per_5min)

p_we <- ggplot(we_avg, aes(x = interval, y = value)) + 
        geom_line() +
        xlab("interval value, seconds")+
        ylab("Average number of steps Weekend")

p_wd <- ggplot(wd_avg, aes(x = interval, y = value)) + 
        geom_line() +
        xlab("interval value, seconds")+
        ylab("Average number of steps Weekday")

library(grid)
library(gtable)
g_we_avg <- ggplotGrob(p_we)
g_wd_avg <- ggplotGrob(p_wd)
g <- rbind(g_we_avg,g_wd_avg,size = "first")

grid.newpage()
grid.draw(g)
```

### 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

I believe this file, CP1.Rmd, satisfies this